%!TEX root =../MacbethThesis.tex
%*******************************************************
% Publications
%*******************************************************
\pdfbookmark[1]{Publications}{publications}
\chapter*{Publications}
Some ideas and figures have appeared previously in the following publications involving the author:

\bigskip

\paragraph{Self-organising management of user-generated data and
knowledge}~\citep{MacbethKER2014}  presents our review of participatory-
sensing applications and the framework for the management of participatory
sensing as a knowledge commons, and according to Ostrom's principles. This
paper covers much of the same ground as \autoref{ch:kc} and \autoref{sec:iad}
of this thesis.

\paragraph{Principled Operationalisation of social systems using
Presage2}~\citep{Macbeth2014}  presents the use of the Presage2 simulation
platform for the purpose of Principled Operationalisation, as per the
sociologically-inspired computing methodology. This is done through the
description of the design and implementation of simulations of a linear public
good game, and the Kyoto Protocol. Some of the figures and description of the
Presage2 platform also appear in \autoref{ch:presage}.

\paragraph{Distributive Justice for Self-Organised Common-Pool Resource
Management}~\citep{Pitt2014}  discusses the application of the theory of
Distributive Justice, as described by Nicholas Rescher, to common-pool
resource management. My contribution was the implementation of the simulation
model, using Presage2, and the collection of results through controlled
experimentation.

\paragraph{Animation of Self-Organising Resource Allocation Using
Presage2}~\citep{Macbeth2012}  is a demonstration and poster, describing the
implementation of a resource-allocation problem using Presage2, and the
animation and visualisation of this problem.

\paragraph{Animation of open multi-agent systems}~\citep{Pitt2011}  presents a
case study of the Presage platform's use for the simulation and animation of
four different multi-agent systems.
